Migration Experience Trajectories: A Three-Mode Principal Component Analysis

### 2022-06-16 preregistration/Preregistration3MPCA.Rmd
Prereg final

---
title           : "Migration Experience Trajectories: A Three Mode Principle Component Analysis"
shorttitle      : "Preregistration: Migration Experience Trajectories"
date            : "`r Sys.setlocale('LC_TIME', 'C'); format(Sys.time(), '%d\\\\. %B %Y')`"
author: 
  - name        : Jannis Kreienkamp
    affiliation : "1"
  - name        : Kai Epstude
    affiliation : "1"
  - name        : Rei Monden
    affiliation : "2"
  - name        : Maximilian Agostini
    affiliation : "1"
  - name        : Peter de Jonge
    affiliation : "1"
  - name        : Laura F. Bringmann
    affiliation : "1"
affiliation:
  - id          : 1
    institution : University of Groningen, Department of Psychology
  - id          : 2
    institution : Hiroshima University, Graduate School of Advanced Science and Engineering
bibliography: "../references.bib"
csl: apa.csl
biblio-style: "apalike" 
output: 
  prereg::cos_prereg:
    citation_package: natbib
    keep_tex:  true
urlcolor: BlueViolet
citecolor: BlueViolet
linkcolor: BlueViolet
editor_options: 
  chunk_output_type: console
header-includes:
   - \usepackage{amsmath}
   - \usepackage{nccmath}
   - \usepackage{enumerate}
   - \usepackage{LatexPackage/trackchanges}
   # - \usepackage[finalnew]{LatexPackage/trackchanges}
   # - \usepackage[finalold]{LatexPackage/trackchanges}
   # - \usepackage[inline]{LatexPackage/trackchanges}
   - \usepackage[usenames,dvipsnames]{xcolor}
   - \addeditor{Jannis}
---

@@

In this manuscript, we aim to assess the utility of a promising analysis technique for describing, summarizing, and understanding the initial developmental data patterns --- \textit{three-mode principle component analysis} (3MPCA). We use 3MPCA specifically because recent studies have laid out the potential effectiveness of dimension reduction procedures, which can address the new forms of person-, variable-, and time point heterogeneity jointly [e.g., @Monden2015]. We analyse data from three experience sampling studies, which followed the migration experiences of recent migrants to the Netherlands. All three studies focus on the psychological adaptation of migrants (i.e., psychological acculturation). However, given that past investigations of psychological acculturation have underexplored the crucial aspect of motivational experiences [@Kreienkamp2022d], the three studies have placed a particular emphasis on the needs, goals, and motives of young migrants. To make full use of all three data sets and to guide future use of dimension reduction procedures with extensive psychological data, we will analyze the data in a descriptive and step wise manner, where we first consider the data jointly and will then assess potential differences between datasets, variables, and time scales.

@@

In this manuscript, we aim to assess the utility of one such promising analysis technique --- \textit{three-mode principle component analysis} (3MPCA). We use 3MPCA specifically because recent studies have laid out the potential effectiveness of dimension reduction procedures, which can address the new forms of person-, variable-, and time point heterogeneity jointly [e.g., @Monden2015]. We are among the first to apply this analysis to experience sampling data and, to the best of our knowledge, we are the first to decompose social psychological experiences. This stands in stark contrast, to a renewed recognition that social psychological phenomena unfold over time, a rapid increase of extensive longitudinal data collections, and a growing interest in understanding the co-development of multiple experience aspects [e.g., @Kreienkamp2022d; @MacInnis2015; @McKeown2017; @Pettigrew2011; @Ward2019].

@@

We collected three experience sampling studies, following the daily migration experiences of young migrants, who had recently arrived in the Netherlands. For all three studies, the data was collected in a three-step procedure:

@@

**Auxiliary variables for multiple imputation**  
For the missing data imputation we follow the procedure reported by @Monden2015, and include any pre-, post-, or daily dairy variables that are significantly ($p$ < .01) and meaningfully correlated ($r$ > .3) with either the key variables or missingness on the key variables. For an overview of the variables that qualify as auxiliary variables see the enclosed data sheets (see `VariableSelectionAnalyses.xlsx'). Additional information about the multiple imputation procedure is also available at [Statistical models] 2.

@@

4. **Three-way ANOVA\*.** We assess the percentages of explained variance for the person-, variable-, and time aspects --- which offers an indication of whether a 3MPCA is useful for the dataset. Based on the grand mean centered variables we conduct a fixed-effects three-way ANOVA of the person-, variable-, and time modes as well as their various interactions. We are particularly interested in the highest order interaction term Person * Variable * Time + error, as a large amount of variance in this effect would speak towards the interdependence of the three modes.

@@

Many of the proposed procedures are descriptive rather then inferential. Importantly, we will use amounts of variances explained as our main fit indices. 
If we perform inferential analyses, we will use the standard p<.05 criteria for determining whether the correlation and regression coefficients are statistically significant. We will use appropriate multiple-comparison corrections whenever multiple tests are performed. We will place particular emphasis on effect sizes (e.g., correlations) in our interpretations of the results wherever possible.

@@

No checks will be performed to determine eligibility for inclusion besides verification that each subject answered each of the variables of interest for a given analysis. Outliers will generally be included in analyses, however we will use sensitivity analyses to assess the robustness of the results to outliers.

>>>>

---
title           : "Migration Experience Trajectories: A Three-Mode Principal Component Analysis"
shorttitle      : "Preregistration: Migration Experience Trajectories"
date            : "`r Sys.setlocale('LC_TIME', 'C'); format(Sys.time(), '%d\\\\. %B %Y')`"
author: 
  - name        : Jannis Kreienkamp
    affiliation : "1"
  - name        : Kai Epstude
    affiliation : "1"
  - name        : Rei Monden
    affiliation : "2"
  - name        : Maximilian Agostini
    affiliation : "1"
  - name        : Peter de Jonge
    affiliation : "1"
  - name        : Laura F. Bringmann
    affiliation : "1"
affiliation:
  - id          : 1
    institution : University of Groningen, Department of Psychology
  - id          : 2
    institution : Hiroshima University, Graduate School of Advanced Science and Engineering
bibliography: "../references.bib"
csl: apa.csl
biblio-style: "apalike" 
output: 
  prereg::cos_prereg:
    citation_package: natbib
    keep_tex:  true
urlcolor: BlueViolet
citecolor: BlueViolet
linkcolor: BlueViolet
editor_options: 
  chunk_output_type: console
header-includes:
   - \usepackage{amsmath}
   - \usepackage{nccmath}
   - \usepackage{enumerate}
   - \usepackage{LatexPackage/trackchanges}
   # - \usepackage[finalnew]{LatexPackage/trackchanges}
   # - \usepackage[finalold]{LatexPackage/trackchanges}
   # - \usepackage[inline]{LatexPackage/trackchanges}
   - \usepackage[usenames,dvipsnames]{xcolor}
   - \addeditor{Jannis}
---

@@

In this manuscript, we aim to assess the utility of a promising analysis technique for describing, summarizing, and understanding the initial developmental data patterns --- \textit{three-mode principal component analysis} (3MPCA). We use 3MPCA specifically because recent studies have laid out the potential effectiveness of dimension reduction procedures, which can address the new forms of person-, variable-, and time point heterogeneity jointly [e.g., @Monden2015]. We analyse data from three experience sampling studies, which followed the migration experiences of recent migrants to the Netherlands. All three studies focus on the psychological adaptation of migrants (i.e., psychological acculturation). However, given that past investigations of psychological acculturation have underexplored the crucial aspect of motivational experiences [@Kreienkamp2022d], the three studies have placed a particular emphasis on the needs, goals, and motives of young migrants. To make full use of all three data sets and to guide future use of dimension reduction procedures with extensive psychological data, we will analyze the data in a descriptive and step wise manner, where we first consider the data jointly and will then assess potential differences between datasets, variables, and time scales.

@@

In this manuscript, we aim to assess the utility of one such promising analysis technique --- \textit{three-mode principal component analysis} (3MPCA). We use 3MPCA specifically because recent studies have laid out the potential effectiveness of dimension reduction procedures, which can address the new forms of person-, variable-, and time point heterogeneity jointly [e.g., @Monden2015]. We are among the first to apply this analysis to experience sampling data and, to the best of our knowledge, we are the first to decompose social psychological experiences. This stands in stark contrast, to a renewed recognition that social psychological phenomena unfold over time, a rapid increase of extensive longitudinal data collections, and a growing interest in understanding the co-development of multiple experience aspects [e.g., @Kreienkamp2022d; @MacInnis2015; @McKeown2017; @Pettigrew2011; @Ward2019].

@@

We collected three experience sampling studies, following the daily migration experiences of young migrants, who had recently arrived in the Netherlands (median time in the Netherlands = 3 Months). For all three studies, the data was collected in a three-step procedure:

@@

**Auxiliary variables for multiple imputation**  
For the missing data imputation we follow the procedure reported by @Monden2015, and include any pre-, post-, or daily dairy variables that are significantly correlated with either the key variables or missingness on the key variables (at $p$ < .01). In case we encounter a convergence problem during the multiple imputation, we only include variable with a correlation larger than .3. For an overview of the variables that qualify as auxiliary variables see the enclosed data sheets (see `VariableSelectionAnalyses.xlsx'). Additional information about the multiple imputation procedure is also available at [Statistical models] 2.

@@

4. **Three-way ANOVA\*.** We assess the percentages of explained variance for the person-, variable-, and time aspects --- which offers an indication of whether a 3MPCA is useful for the dataset. Based on the grand mean centered variables we conduct a fixed-effects three-way ANOVA of the person-, variable-, and time modes as well as their various interactions. We are particularly interested in the highest order interaction term Person * Variable * Time + error, as a large amount of variance in this effect would speak towards the possible interdependence of the three modes.

@@

Many of the proposed procedures are descriptive rather than inferential. Importantly, we will use amounts of variances explained as our main fit indices. 
If we perform inferential analyses, we will use the standard p<.05 criteria for determining whether the correlation and regression coefficients are statistically significant. We will use appropriate multiple-comparison corrections whenever multiple tests are performed (such as Bonferroni correction). We will place particular emphasis on effect sizes (e.g., correlations) in our interpretations of the results wherever possible.

@@

No checks will be performed to determine eligibility for inclusion besides verification that each subject answered each of the variables of interest for a given analysis. Outliers will generally be included in analyses. However, we will use sensitivity analyses to assess the robustness of the results to outliers.

### 2022-06-16 preregistration/Preregistration3MPCA.tex
Prereg final

\documentclass[]{article}
    \usepackage{lmodern}
    \usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
  \else % if luatex or xelatex
\ifxetex
\usepackage{mathspec}
\usepackage{xltxtra,xunicode}
\else
  \usepackage{fontspec}
\fi
\defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
\newcommand{\euro}{€}
        \fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
  \usepackage{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
  \usepackage[left=2.5in,bottom=1.25in,top=1.25in,right=1in]{geometry}
  \ifxetex
\usepackage[setpagesize=false, % page size defined by xetex
            unicode=false, % unicode breaks when used with xetex
            xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
bookmarks=true,
pdfauthor={},
pdftitle={Migration Experience Trajectories: A Three Mode Principle Component Analysis},
colorlinks=true,
citecolor=BlueViolet,
urlcolor=BlueViolet,
linkcolor=BlueViolet,
pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{apalike}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

@@

\setlength{\droptitle}{-2em}
\title{Migration Experience Trajectories: A Three Mode Principle
Component Analysis}
\pretitle{\begin{changemargin}{-8pc}{0pc} \centering\large Preregistration\\ \Huge}
\posttitle{\end{changemargin}}
  \author{
          Jannis Kreienkamp\textsuperscript{1},
          Kai Epstude\textsuperscript{1},
          Rei Monden\textsuperscript{2},
          Maximilian Agostini\textsuperscript{1},
          Peter de Jonge\textsuperscript{1},
          Laura F.
Bringmann\textsuperscript{1}          \\ \vspace{0.5cm}
              \textsuperscript{1} University of Groningen, Department of
Psychology\\
              \textsuperscript{2} Hiroshima University, Graduate School
of Advanced Science and Engineering      }

  \def\affdep{{"", "", "", "", "", ""}}%
  \def\affcity{{"", "", "", "", "", ""}}%
  \preauthor{\begin{changemargin}{-8pc}{0pc} \centering\large}
  \postauthor{\end{changemargin}}
\date{28. May 2022}
\predate{\begin{changemargin}{-8pc}{0pc} \centering\large\emph}
\postdate{\end{changemargin}}
\usepackage{fancyhdr}
\pagestyle{fancy}
\setlength{\headheight}{14.0pt}
\renewcommand{\headrulewidth}{0pt}
\lhead{}
\rhead{\large\textsc{\MakeLowercase{Preregistration: Migration
Experience Trajectories}}}

@@

Migration Experience Trajectories: A Three Mode Principle Component
Analysis

@@

In this manuscript, we aim to assess the utility of a promising analysis
technique for describing, summarizing, and understanding the initial
developmental data patterns ---
\textit{three-mode principle component analysis} (3MPCA). We use 3MPCA
specifically because recent studies have laid out the potential
effectiveness of dimension reduction procedures, which can address the
new forms of person-, variable-, and time point heterogeneity jointly
\citep[e.g.,][]{Monden2015}. We analyse data from three experience
sampling studies, which followed the migration experiences of recent
migrants to the Netherlands. All three studies focus on the
psychological adaptation of migrants (i.e., psychological
acculturation). However, given that past investigations of psychological
acculturation have underexplored the crucial aspect of motivational
experiences \citep{Kreienkamp2022d}, the three studies have placed a
particular emphasis on the needs, goals, and motives of young migrants.
To make full use of all three data sets and to guide future use of
dimension reduction procedures with extensive psychological data, we
will analyze the data in a descriptive and step wise manner, where we
first consider the data jointly and will then assess potential
differences between datasets, variables, and time scales.

@@

In this manuscript, we aim to assess the utility of one such promising
analysis technique --- \textit{three-mode principle component analysis}
(3MPCA). We use 3MPCA specifically because recent studies have laid out
the potential effectiveness of dimension reduction procedures, which can
address the new forms of person-, variable-, and time point
heterogeneity jointly \citep[e.g.,][]{Monden2015}. We are among the
first to apply this analysis to experience sampling data and, to the
best of our knowledge, we are the first to decompose social
psychological experiences. This stands in stark contrast, to a renewed
recognition that social psychological phenomena unfold over time, a
rapid increase of extensive longitudinal data collections, and a growing
interest in understanding the co-development of multiple experience
aspects
\citep[e.g.,][]{Kreienkamp2022d, MacInnis2015, McKeown2017, Pettigrew2011, Ward2019}.

@@

\textbf{Auxiliary variables for multiple imputation}\\
For the missing data imputation we follow the procedure reported by
\citet{Monden2015}, and include any pre-, post-, or daily dairy
variables that are significantly (\(p\) \textless{} .01) and
meaningfully correlated (\(r\) \textgreater{} .3) with either the key
variables or missingness on the key variables. For an overview of the
variables that qualify as auxiliary variables see the enclosed data
sheets (see `VariableSelectionAnalyses.xlsx'). Additional information
about the multiple imputation procedure is also available at
\protect\hyperlink{statistical-models}{Statistical models} 2.

@@

  amount of variance in this effect would speak towards the

@@

Many of the proposed procedures are descriptive rather then inferential.
Importantly, we will use amounts of variances explained as our main fit
indices. If we perform inferential analyses, we will use the standard
p\textless.05 criteria for determining whether the correlation and
regression coefficients are statistically significant. We will use
appropriate multiple-comparison corrections whenever multiple tests are
performed. We will place particular emphasis on effect sizes (e.g.,
correlations) in our interpretations of the results wherever possible.

>>>>

\documentclass[]{article}
    \usepackage{lmodern}
    \usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
  \else % if luatex or xelatex
\ifxetex
\usepackage{mathspec}
\usepackage{xltxtra,xunicode}
\else
  \usepackage{fontspec}
\fi
\defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
\newcommand{\euro}{€}
        \fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
  \usepackage{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
  \usepackage[left=2.5in,bottom=1.25in,top=1.25in,right=1in]{geometry}
  \ifxetex
\usepackage[setpagesize=false, % page size defined by xetex
            unicode=false, % unicode breaks when used with xetex
            xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
bookmarks=true,
pdfauthor={},
pdftitle={Migration Experience Trajectories: A Three-Mode Principal Component Analysis},
colorlinks=true,
citecolor=BlueViolet,
urlcolor=BlueViolet,
linkcolor=BlueViolet,
pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{apalike}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

@@

\setlength{\droptitle}{-2em}
\title{Migration Experience Trajectories: A Three-Mode Principal
Component Analysis}
\pretitle{\begin{changemargin}{-8pc}{0pc} \centering\large Preregistration\\ \Huge}
\posttitle{\end{changemargin}}
  \author{
          Jannis Kreienkamp\textsuperscript{1},
          Kai Epstude\textsuperscript{1},
          Rei Monden\textsuperscript{2},
          Maximilian Agostini\textsuperscript{1},
          Peter de Jonge\textsuperscript{1},
          Laura F.
Bringmann\textsuperscript{1}          \\ \vspace{0.5cm}
              \textsuperscript{1} University of Groningen, Department of
Psychology\\
              \textsuperscript{2} Hiroshima University, Graduate School
of Advanced Science and Engineering      }

  \def\affdep{{"", "", "", "", "", ""}}%
  \def\affcity{{"", "", "", "", "", ""}}%
  \preauthor{\begin{changemargin}{-8pc}{0pc} \centering\large}
  \postauthor{\end{changemargin}}
\date{01. June 2022}
\predate{\begin{changemargin}{-8pc}{0pc} \centering\large\emph}
\postdate{\end{changemargin}}
\usepackage{fancyhdr}
\pagestyle{fancy}
\setlength{\headheight}{14.0pt}
\renewcommand{\headrulewidth}{0pt}
\lhead{}
\rhead{\large\textsc{\MakeLowercase{Preregistration: Migration
Experience Trajectories}}}

@@

Migration Experience Trajectories: A Three-Mode Principal Component
Analysis

@@

In this manuscript, we aim to assess the utility of a promising analysis
technique for describing, summarizing, and understanding the initial
developmental data patterns ---
\textit{three-mode principal component analysis} (3MPCA). We use 3MPCA
specifically because recent studies have laid out the potential
effectiveness of dimension reduction procedures, which can address the
new forms of person-, variable-, and time point heterogeneity jointly
\citep[e.g.,][]{Monden2015}. We analyse data from three experience
sampling studies, which followed the migration experiences of recent
migrants to the Netherlands. All three studies focus on the
psychological adaptation of migrants (i.e., psychological
acculturation). However, given that past investigations of psychological
acculturation have underexplored the crucial aspect of motivational
experiences \citep{Kreienkamp2022d}, the three studies have placed a
particular emphasis on the needs, goals, and motives of young migrants.
To make full use of all three data sets and to guide future use of
dimension reduction procedures with extensive psychological data, we
will analyze the data in a descriptive and step wise manner, where we
first consider the data jointly and will then assess potential
differences between datasets, variables, and time scales.

@@

In this manuscript, we aim to assess the utility of one such promising
analysis technique --- \textit{three-mode principal component analysis}
(3MPCA). We use 3MPCA specifically because recent studies have laid out
the potential effectiveness of dimension reduction procedures, which can
address the new forms of person-, variable-, and time point
heterogeneity jointly \citep[e.g.,][]{Monden2015}. We are among the
first to apply this analysis to experience sampling data and, to the
best of our knowledge, we are the first to decompose social
psychological experiences. This stands in stark contrast, to a renewed
recognition that social psychological phenomena unfold over time, a
rapid increase of extensive longitudinal data collections, and a growing
interest in understanding the co-development of multiple experience
aspects
\citep[e.g.,][]{Kreienkamp2022d, MacInnis2015, McKeown2017, Pettigrew2011, Ward2019}.

@@

\textbf{Auxiliary variables for multiple imputation}\\
For the missing data imputation we follow the procedure reported by
\citet{Monden2015}, and include any pre-, post-, or daily dairy
variables that are significantly correlated with either the key
variables or missingness on the key variables (at \(p\) \textless{}
.01). In case we encounter a convergence problem during the multiple
imputation, we only include variable with a correlation larger than .3.
For an overview of the variables that qualify as auxiliary variables see
the enclosed data sheets (see `VariableSelectionAnalyses.xlsx').
Additional information about the multiple imputation procedure is also
available at \protect\hyperlink{statistical-models}{Statistical models}
2.

@@

  amount of variance in this effect would speak towards the possible

@@

Many of the proposed procedures are descriptive rather than inferential.
Importantly, we will use amounts of variances explained as our main fit
indices. If we perform inferential analyses, we will use the standard
p\textless.05 criteria for determining whether the correlation and
regression coefficients are statistically significant. We will use
appropriate multiple-comparison corrections whenever multiple tests are
performed (such as Bonferroni correction). We will place particular
emphasis on effect sizes (e.g., correlations) in our interpretations of
the results wherever possible.
